logbook simulations

### 2020-03-23 manuscript/drafts/manuscript.Rmd
- add workflow optimization hyperparameters - add log hyperparameter optimization - add first simulation results BCTW model

Now there are a few technical details. Many different versions of such algorithms exist. Many of such algorithms have been described in the active learning literature and have been applied in the systematic reviewing process.  Not exhaustive, but the algorithm can apply many different strategies to arrive at its predictions, which can be divided in following parameters: classifier, feature extraction strategy, balancing, query strategy.

@@

#### Word representation 
To be able to predict whether a paper needs to be included or excluded (e.g. to predict class), 
the classifier needs some features from the papers.  
As features we use the title and abstract from every paper - this is what is prescribed by prisma, (check!) 
The classifier cannot predict the paper class from the raw titles and abstracts as they are.
Therefore, the content of the texts needs to be transformed into numerical representations called feature vectors.

A classical example is a 'bag of words' representation. 
For each each text, the number of occurrences of each word is stored. 
This leads to n features, where n is the number of distinct words in the texts [@scikit-learn].
The bag-of-words method is simplistic and will highly value often occuring but otherwise meaningless words such as "and". 
Term-frequency Inverse Document Frequency (TF-IDF) [@Ramos2003] circumvents this problem by adjusting a term frequency in a text with the inverse docuement frequency, the frequency of a given word in the entire corpus.

>>>>

Now there are a few technical details. Many different versions of such algorithms exist. Many of such algorithms have been described in the active learning literature and have been applied in the systematic reviewing process.  Not exhaustive, but the algorithm can apply many different strategies to arrive at its predictions, which can be divided in following parameters: classifier, feature extraction strategy, balancing, query strateg y.

@@

#### Word representation 
To predict the class of a publication class (e.g. whether a publication should be included or excluded), the classifier uses information from the publications in the dataset. 
Examples of such information are titles and abstracts. 
However, the classifier cannot predict the publication class from titles and abstracts as they are. 
Their textual content to needs to be mapped to feature vectors. 
This process of numerically representing textual content is called 'word embeddings'.

A classical example of word embeddings a 'bag of words' representation. 
For each text in the data set, the number of occurrences of each word is stored. 
This leads to n features, where n is the number of distinct words in the texts [@scikit-learn].
The bag-of-words method is simplistic and will highly value often occuring but otherwise meaningless words such as "and". 
Term-frequency Inverse Document Frequency (TF-IDF) [@Ramos2003] circumvents this problem by adjusting a term frequency in a text with the inverse docuement frequency, the frequency of a given word in the entire data set.

The model expedits features of previously labeled publications and 
Based on features of the previously labeled inclusions and exclusions, the model

The features used are title and abstract from every publication. - this is what is prescribed by prisma, (check!)

### 2020-03-23 simulation/README.Rmd
- add workflow optimization hyperparameters - add log hyperparameter optimization - add first simulation results BCTW model

This is the log of simulations. 
There are three stages, we start with the first where we are comparing different classifiers.

```{r include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r, echo = FALSE}
conditiongrid <- readRDS("00_conditiongrid.RDS")

conditiongrid$condition %>%
  kable(format = "markdown") %>%
  pack_rows("Classifier", 1, 5) %>%
  pack_rows("Feature Extraction", 6, 10) %>%
  pack_rows("Balance Strategy", 11, 20)
```

### Version information
Personal device

Cartesius supercomputer

@@

hyperparametersets %>%
  kable(format = "markdown")
```

@@

## 1.1 Optimizing (hyper)parameters per model, per dataset (1-1)

### 1.1.1 Ace
```{r}
lubridate::date("2020-03-17")
```

Optimization run on own device 
```{bash, eval = FALSE}
# RCTW
mpirun -n 2 asreview hyper-active -m nb -b double -e tfidf -q max -d ace -n 10000 -r 8 --mpi 
```

Optimization run on Cartesius
```{bash, eval = FALSE}
# RCTW
./cart_hyper_run.py hyper-active -t 10 -m rf -b double -e tfidf -q max -d ace

@@

### 1.1.2 Nudging
### 1.1.3 PTSD
### 1.1.4 Software
### 1.1.5 Wilson

# Stage 2 - Feature exraction 
Comparing tfidf with doc2vec

# Stage 3 - Balance strategy
Redo all the analyses with a different balance strategy (aggressive undersampling?)
Probably won't have time for this.

>>>>

This is the log of simulations.

# Simulations

## Hyperparameter optimization
There are three stages, we start with the first where we are comparing different classifiers.

```{r include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
datasets <- c("ace", "ptsd", "nudging", "software", "wilson")
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
conditiongrid <- readRDS("00_conditiongrid.RDS")

conditiongrid %>%
  kable(format = "markdown")
```

### Version information
Personal device: macos catalina 10.15.2

Cartesius supercomputer 
asreview 0.7.2
asreview simulation branch reorganization.

@@

# to-do grid 
todo <- hyperparametersets

# set emojis for grid 
check <- function(todo, m, ho, d){
  todo[todo$model %in% m & todo$hyperopt == ho, d] <- emo::ji("check") 
  return(todo)
}

busy <- function(todo, m, ho, d){
  todo[todo$model %in% m & todo$hyperopt == ho, d] <- emo::ji("timer") 
  return(todo)
}

huh <- function(todo, m, ho, d){
  todo[todo$model %in% m & todo$hyperopt == ho, d] <- emo::ji("question") 
  return(todo)
}

not <- function(todo, m, ho, d){
  todo[todo$model %in% m & todo$hyperopt == ho, d] <- emo::ji("x") 
  return(todo)
}

```

@@

## 1.1 Optimizing (hyper)parameters per model, per dataset (1-1)
```{r}
lubridate::date("2020-03-17")
# i am working on svm model for all datasets 
todo <- busy(todo, "SCTW", "one", "ace")
```

Optimization run on Cartesius
```{bash, eval = FALSE}
# BCTW

# RCTW

# SCTW 
./cart_hyper_run.py hyper-active -t 12:00:00 -m svm -b double -e tfidf -q max -d ace -r 23 # job id 7768402

***

```{r}
lubridate::date("2020-03-18")
todo <- check(todo, "BCTW", "one", datasets)
todo <- check(todo, "SCTW", "one", "ace")
```

Store config files from job 7768402, SCTW model.

```{bash, eval = FALSE}
asreview create-config output/active/svm_max_double_tfidf/ace/trials.pkl -o config/SCTW_ace.ini
```

Transfer config files to my computer home directory 
```{bash, eval = FALSE}
# in own device. 
scp gerbrich@cartesius.surfsara.nl:asreview/config/SCTW_ace.ini ~
```

```{bash, eval = FALSE}
# BCTW
./cart_hyper_run.py hyper-active -t 03:00:00 -m nb -b double -e tfidf -q max -d ace -r 23 # 7772446
./cart_hyper_run.py hyper-active -t 03:00:00 -m nb -b double -e tfidf -q max -d nudging -r 23 # 7772450
./cart_hyper_run.py hyper-active -t 03:00:00 -m nb -b double -e tfidf -q max -d ptsd -r 23 # 7772452
./cart_hyper_run.py hyper-active -t 03:00:00 -m nb -b double -e tfidf -q max -d software -r 23 # 7772454
./cart_hyper_run.py hyper-active -t 03:00:00 -m nb -b double -e tfidf -q max -d wilson -r 23 # 7772456
# extract all trials.pkl files 
scp gerbrich@cartesius.surfsara.nl:asreview/output/active/nb_max_double_tfidf/ace/trials.pkl ~/hyperopt/trials/BCTW_ace.pkl 
scp gerbrich@cartesius.surfsara.nl:asreview/output/active/nb_max_double_tfidf/nudging/trials.pkl ~/hyperopt/trials/BCTW_nudging.pkl 
scp gerbrich@cartesius.surfsara.nl:asreview/output/active/nb_max_double_tfidf/ptsd/trials.pkl ~/hyperopt/trials/BCTW_ptsd.pkl
scp gerbrich@cartesius.surfsara.nl:asreview/output/active/nb_max_double_tfidf/software/trials.pkl ~/hyperopt/trials/BCTW_software.pkl 
scp gerbrich@cartesius.surfsara.nl:asreview/output/active/nb_max_double_tfidf/wilson/trials.pkl ~/hyperopt/trials/BCTW_wilson.pkl 
# create config files
asreview create-config trials/BCTW_ace.pkl -o config/BCTW_ace.ini
asreview create-config trials/BCTW_nudging.pkl -o config/BCTW_nudging.ini
asreview create-config trials/BCTW_ptsd.pkl -o config/BCTW_ptsd.ini
asreview create-config trials/BCTW_software.pkl -o config/BCTW_software.ini
asreview create-config trials/BCTW_wilson.pkl -o config/BCTW_wilson.ini

# RCTW

@@

# NCTW

```

***

```{r}
lubridate::date("2020-03-19")

#todo <- check(todo, "BCTW", "one", datasets)
```

### simulations with nb 
```{bash, eval = FALSE}
# create config files within cartesius (scripts folder)
asreview create-config output/active/nb_max_double_tfidf/ace/trials.pkl -o config/BCTW_ace.ini
asreview create-config output/active/nb_max_double_tfidf/nudging/trials.pkl  -o config/BCTW_nudging.ini
asreview create-config output/active/nb_max_double_tfidf/ptsd/trials.pkl  -o config/BCTW_ptsd.ini
asreview create-config output/active/nb_max_double_tfidf/software/trials.pkl  -o config/BCTW_software.ini
asreview create-config output/active/nb_max_double_tfidf/wilson/trials.pkl  -o config/BCTW_wilson.ini
```

```{bash, eval = FALSE}
# run simulations within cartesius
./cart_simulate.py data/ace.csv --config_file config/BCTW_ace.ini --state_file "simoutput/bctw/ace/BCTW_ace.h5" -r 23

./cart_simulate.py data/nudging.csv --config_file config/BCTW_nudging.ini --state_file "simoutput/bctw/nudging/BCTW_nudging.h5" -r 23

./cart_simulate.py data/ptsd.csv --config_file config/BCTW_ptsd.ini --state_file "simoutput/bctw/ptsd/BCTW_ptsd.h5" -r 23

./cart_simulate.py data/software.csv --config_file config/BCTW_software.ini --state_file "simoutput/bctw/software/BCTW_software.h5" -r 23

./cart_simulate.py data/wilson.csv --config_file config/BCTW_wilson.ini --state_file "simoutput/bctw/wilson/BCTW_wilson.h5" -r 23 
```

### extract simulation results nb 
```{bash}
# transfer files to device 
scp -r gerbrich@cartesius.surfsara.nl:asreview/simoutput /Users/gerbrich/Documents/MSBBSS/asreviewthesis/asreview-thesis/simulation_output
```

```{bash}
# on own device
asreview plot asreview-thesis/simulation_output
```

***
```{r}
lubridate::date("2020-03-22")
todo <- busy(todo, "SCTW", "one", datasets[2:5])
todo <- busy(todo, c("LCTW", "RCTW", "NCDW"), "one", "ace")
todo <- not(todo, "NCTW", "one", datasets)
# do the remaining datasets of svm, and one test run for L and one for R. 
```

Sent hyperparameters for optimization
```{bash, eval = FALSE}
# SCTW
./cart_hyper_run.py hyper-active -t 12:00:00 -m svm -b double -e tfidf -q max -d nudging -r 23 # 7805983
./cart_hyper_run.py hyper-active -t 12:00:00 -m svm -b double -e tfidf -q max -d ptsd -r 23 # 7805984
./cart_hyper_run.py hyper-active -t 12:00:00 -m svm -b double -e tfidf -q max -d software -r 23 # 7805985
./cart_hyper_run.py hyper-active -t 12:00:00 -m svm -b double -e tfidf -q max -d wilson -r 23 # 7805986

# LCTW 
./cart_hyper_run.py hyper-active -t 12:00:00 -m lr -b double -e tfidf -q max -d ace -r 23 # 7805987

# RCTW 
./cart_hyper_run.py hyper-active -t 12:00:00 -m rf -b double -e tfidf -q max -d ace -r 23 # 7805989

# NCTW does not work well with tfidf, save for second round with doc2vec: 
# NCDW
./cart_hyper_run.py hyper-active -t 12:00:00 -m nn-2-layer -b double -e doc2vec -q max -d ace -r 23 # 7806021

```

***
```{r}
lubridate::date("2020-03-23")
todo <- check(todo, "SCTW", "one", datasets[3:5])
```

Resimulate BCTW models.
 - create-config file from trials.pkl: n_instances 20, 1/1
 - how to overrule this parameters? (not possible through cli)

```{bash, eval = FALSE}
./cart_simulate.py data/ace.csv --config_file config/BCTW_ace.ini --state_file "simoutput/bctw/ace/results_ace.h5" -r 23 
./cart_simulate.py data/nudging.csv --config_file config/BCTW_nudging.ini --state_file "simoutput/bctw/nudging/results_nudging.h5" -r 23 
./cart_simulate.py data/ptsd.csv --config_file config/BCTW_ptsd.ini --state_file "simoutput/bctw/ptsd/results_ptsd.h5" -r 23 
./cart_simulate.py data/software.csv --config_file config/BCTW_software.ini --state_file "simoutput/bctw/software/results_software.h5" -r 23 
./cart_simulate.py data/wilson.csv --config_file config/BCTW_wilson.ini --state_file "simoutput/bctw/wilson/results_wilson.h5" -r 23 
```

Process BCTW models into table
 - python script to extract statistics
 -

Checkout hyperopt SCTW
Nudging dataset doesn't optimize well? 
```{r}
todo <- huh(todo, "SCTW", "one", "nudging")
```

Checkout results first hyperopt: 
LCTW - no output?? 
RCTW - only 'current'
```{r}
todo <- huh(todo, c("LCTW", "RCTW"), "one", "ace")
```

NCDW - loss still going down, sent in for another 12 hours. 
```{bash, eval = FALSE}
./cart_hyper_run.py hyper-active -t 12:00:00 -m nn-2-layer -b double -e doc2vec -q max -d ace -r 23 # 7811926
```

(re)Simulate BCTW models
Simulate SCTW models

***

```{r}
lubridate::date("2020-03-24")
# todo <- busy(todo, "SCTW", "one", datasets[2:5])
```

Process SCTW models into table

Simulate LCTW, RCTW and NCTW models

Create a nice plot

Send first version of table

***

```{r}
# save the to-do grid in markdown style
todo[todo==0] <- emo::ji("white square button")
todo %>% kable(format = "markdown")

saveRDS(todo %>% kable(format = "markdown"), 
        file = "02_todo_hp.RDS")
```

## Stage 2 - Feature extraction 
Comparing tfidf with doc2vec

## Stage 3 - Balance strategy
Redo all the analyses with a different balance strategy (aggressive undersampling?)
Probably won't have time for this.

# The data 
On 17 march 2020:

```
************  ace.csv  ************

Number of papers:            2235
Number of inclusions:        41 (1.83%)
Number of exclusions:        2194 (98.17%)
Number of unlabeled:         0 (0.00%)
Average title length:        123
Average abstract length:     1623
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  nudging.csv  ************

Number of papers:            1847
Number of inclusions:        100 (5.41%)
Number of exclusions:        1747 (94.59%)
Number of unlabeled:         0 (0.00%)
Average title length:        109
Average abstract length:     1831
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  ptsd.csv  ************

Number of papers:            5031
Number of inclusions:        38 (0.76%)
Number of exclusions:        4993 (99.24%)
Number of unlabeled:         0 (0.00%)
Average title length:        104
Average abstract length:     1537
Number of missing titles:    1 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  software.csv  ************

Number of papers:            8896
Number of inclusions:        104 (1.17%)
Number of exclusions:        8792 (98.83%)
Number of unlabeled:         0 (0.00%)
Average title length:        81
Average abstract length:     896
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  wilson.csv  ************

Number of papers:            2333
Number of inclusions:        23 (0.99%)
Number of exclusions:        2310 (99.01%)
Number of unlabeled:         0 (0.00%)
Average title length:        83
Average abstract length:     1325
Number of missing titles:    1 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)
```
